---
title: "Quarto Exercise 6"
author: "Jacob Green"
format: html
---

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(reactable)
library(crosstalk)
library(plotly)

universities <- read.csv("https://manctsui.github.io/datasets/timesData.csv")
```

## Problem 1A

Find the three countries with the most universities in the rankings.

```{r}
top_countries <- universities |>
  count(country, name = "number_of_universities") |>
  arrange(desc(number_of_universities)) |>
  slice_head(n = 3)

reactable(top_countries)
```

## Problem 1B

Display the result from Problem 1A as a bar chart.

```{r}
ggplot(top_countries,
       aes(x = reorder(country, number_of_universities),
           y = number_of_universities)) +
  geom_col() +
  coord_flip() +
  labs(
    x = "Country",
    y = "Number of Universities",
    title = "Countries with the Most Universities in the Rankings"
  )
```

## Problem 2A

Display the top 50 universities in 2011 with only rank and university name.

```{r}
top_50_2011 <- universities |>
  filter(year == 2011) |>
  mutate(rank = as.numeric(gsub("=", "", world_rank))) |>
  arrange(rank) |>
  slice_head(n = 50) |>
  select(rank, university_name)

reactable(top_50_2011)
```

## Problem 2B

The rank column would not be the best key for a SharedData object because a key should uniquely identify each row. Rankings can contain ties, so the same rank could appear more than once. The university name is a better choice because each university in this table is a separate observation.

## Problem 2C

Create a slider that filters universities by rank.

```{r}
shared_top_50 <- SharedData$new(
  top_50_2011,
  key = ~university_name
)

filter_slider(
  id = "rank_filter",
  label = "University Rank",
  sharedData = shared_top_50,
  column = ~rank
)

reactable(shared_top_50)
```

## Problem 3

Create two linked scatterplots using the 2016 data.

```{r}
universities_2016 <- universities |>
  filter(year == 2016) |>
  mutate(
    num_students = as.numeric(gsub(",", "", num_students)),
    student_staff_ratio = as.numeric(student_staff_ratio),
    teaching = as.numeric(teaching)
  ) |>
  drop_na(num_students, student_staff_ratio, teaching)

shared_2016 <- SharedData$new(
  universities_2016,
  key = ~university_name,
  group = "universities"
)
```

### Number of Students vs. Student-Staff Ratio

```{r}
plot_ly(
  shared_2016,
  x = ~num_students,
  y = ~student_staff_ratio,
  type = "scatter",
  mode = "markers",
  text = ~university_name
) |>
  layout(
    xaxis = list(title = "Number of Students"),
    yaxis = list(title = "Student-Staff Ratio")
  )
```

### Student-Staff Ratio vs. Teaching Score

```{r}
plot_ly(
  shared_2016,
  x = ~student_staff_ratio,
  y = ~teaching,
  type = "scatter",
  mode = "markers",
  text = ~university_name
) |>
  layout(
    xaxis = list(title = "Student-Staff Ratio"),
    yaxis = list(title = "Teaching Score")
  )
```


